SAT Competition 2020\tnoteref{title}

### 2021-01-26 paper/main.tex
fix tabel reference

\subsection{Planning Instances}
Classical planning is the problem of finding a sequence of actions -- a plan --
that transforms the world from some initial state to a goal state. In 1992 Kautz
and Selman
\cite{Kautz1992} proposed to encode planning as satisfiability. In their
encoding the problem of finding a plan of length $i$ (\textit{i.e.,} the
\emph{makespan}) is translated into a Boolean formula $F_i$ that is satisfiable
exactly if a plan of length $i$ \emph{or less} exists. In later encodings
multiple actions can be executed \emph{in parallel} allowing longer plans to be
found by solving smaller formulas \cite{Rintanen2006, Rintanen2007, Balyo2013}.

Finding the minimal makespan $i$ for which $F_i$ is satisfiable is important for
SAT-based planning in general and the generation of this benchmark set in
particular. The minimal makespan depends on the planning task and the used
encoding. The hardest formulas that a SAT-based planner has to solve are usually
the last unsatisfiable $F_i$ before the next higher makespan becomes satisfiable
\cite{Rintanen2006}. Therefore we preferably pick the last unsatisfiable
makespan for each planning task to generate unsatisfiable instances. For
planning tasks where this makespan cannot be determined with available
computational resources, we use a \emph{sequential} encoding, where the minimal
makespan equals the length of the shortest valid plan. Together with known
bounds\footnote{Bounds on plan length are available for some planning tasks from
  the \emph{optimal track} that have \emph{unit cost} actions.} on the optimal
plan length we can generate SAT formulas with predetermined satisfiability for
hard planning problems.

@@

@@

% - 86 out of 100 biggest (number of clauses) are planning
% - 109 / 200 in the planning track are not solved by any solver
    % - 186 / 600 in total
% - all 200 above 90\% binary clauses and half have above 99%
The instances of the Planning track are big compared to the rest of the
main track. Using the number of clauses as a metric; out of the $100$ biggest
instances
% of the main-track
$86$ belong to the Planning track. The large size can mainly be attributed to
binary clauses. On average, more than $98$\% of the clauses are binary for
planning instances, where the average for the rest of the instances from the
Main track is below $60$\%.

Despite the large size, only a few instances directly caused memory problems for
more than just a few solvers. In most cases the \SI{128}{\giga\byte} of RAM
available were sufficient.
% The saving grace might have been
% This can be attributed to
% the \SI{128}{\giga\byte} of RAM available to each solver.
However, the three biggest instances from the \emph{childsnack} family of
planning problems encoded with the \emph{hierarchical} planner \emph{Tree-REX}
caused over half of the solvers to abort with some error message related to
memory. Other solvers were still able to solve the same instances within $300$
seconds.

@@

A greedy set cover of the solved instances by solvers of the Main track is presented in Table~\ref{tab:gscMain}. 
In the first iteration, the solver which solved the highest number of instances is selected; in our case it was \solver{Kissat-sat} with 264 instances as we know already from Table~\ref{tab:mainALL}.
Given those 264 instances as already covered, \solver{CaDiCaL-alluip} was the best in further contributing to the set by additional $22$ instances in iteration $2$.
We can see that the further iterations tend to add very little,
with the final four iterations adding one instance each. Note that each solver
that managed to solve an instance uniquely (i.e.~being the only solver to do so)
is showing up in the table. This is, for example the case of the last listed  
\solver{exp-V-LGB-MLD-CBT-DL}.
Highlighting unique solutions is a property of the greedy set cover metric
which we believe is congruent with documenting contributions to the state of the art.

@@

\subsection{Similarity of Solvers}
To investigate how similar the solvers of the Main track are to each other, we
define a similarity metric based on the measured runtimes. For that we first
remove $186$ benchmarks from the collected data that have not been solved by any
solver. For the rest of the benchmarks we use the PAR-2 score,
% based on the \num{5000}-second timeout
i.\,e., for each solver we set a score of \num{10000} for unsolved instances.
This leaves us with $414$ floats
$S_{1},\dots,S_{n}$ for each solver $S$. The similarity for two solvers $A$
and $B$, normalized to the interval $[0,1]$ is defined
% as:
by Equation~\ref{eq-similarity}.

\begin{equation}
\label{eq-similarity}
	similarity(A, B) =\quad 1 - \frac{\sum{|A_{i} - B_{i}|}}{414 \cdot 10000}
\end{equation}
We calculated the \emph{similarity} between the $30$ solvers which achieved the
best PAR-2 score in the main track. The results are displayed in
Figure~\ref{fig-similarity-main} as a \emph{heat map}. Additionally the result
of hierarchically clustering the solvers based on their similarity is
illustrated as a \emph{dendrogram}. The height at which two solvers or clusters
are joined reflects how similar they are. For example
enabling the \emph{trail saving} in \emph{CaDiCaL-alluip} has no impact on the
runtime, resulting in a similarity above \num{0.999}. Thus the two solvers are joined very low in the dendrogram.
% A similar visualization was used in the SAT Challenge 2012 \todo{cite}. They used

One big cluster we can identify are the \emph{Maple} decedents in the lower left
corner, all of which are modifications of the winners of the SAT Competition
2018 and the SAT Race 2019. 
%The solver \emph{mergesat} does not quite fit into the cluster and has a comparatively high similarity to the \emph{CryptoMiniSat} (\emph{CMS}) family at the top right. With the exception of \emph{mergesat} and maybe \emph{DurianSat} we observe a high similarity within the cluster.

The three \emph{Relaxed} solvers by \emph{Zhang} and \emph{Cai} are using the
same code base as a lot of the solvers in the Maple-Cluster but their overall
performance is better and closer to the CMS-Cluster.

The CMS versions differ in their implementation of stochastic local search (SLS)
but overall have a similar performance.

The last big cluster in the center of the heat map is comprised of the solvers
originally written by Biere. The three configurations of \emph{Kissat} form a
sub-cluster, even though the configurations are fairly distinct from each other.
% , specialized for satisfiable and unsatisfiable instances.
% One targets satisfiable instances, while the other is focused on unsatisfiable
% instances and a third alternating between them.

The other sub-cluster consists of
modifications to the 2019 version of \emph{\mbox{CaDiCaL}}. The new version
CaDiCaL-2020 has a worse overall performance than
% has an overall worse per
% is lacking behind
the other solvers in this cluster.

\begin{figure}[tp]
    \centering
    \includegraphics[width=\linewidth]{similarity/cross.pdf}
    \caption{Heat-map and dendrogram (top) based on the runtime similarity of the solvers participating in the main track.
      % Similarity is defined as
      % one minus the average taxicab distance normalized to the interval [0,1].
      % Spearman's rank correlation coefficient.
      Darker regions mean that the solvers are more similar.
      A more precise relation between color and similarity-value together with a histogram of the values that appear is given at the bottom.
    }
    \label{fig-similarity-main}
\end{figure}

% To estimate the similarity between the performance of the solvers of the Main
% Track we define a similarity metric $s$.
% identify each solver with a point in the 600-dimensional real vector
% space. Each component describes the PAR-2 runtime on one instance. We map the
% taxicab distance to the $[0, 1]$ interval and then to determine how similar two
% solvers are.

\subsection{Sampling}
\includegraphics[width=\textwidth]{stability/ALL_random_smpling_correlations.png}
\includegraphics[width=\textwidth]{stability/fam_leave_one_out_corr.png}

>>>>

\subsection{Planning Instances}
Classical planning is the problem of finding a sequence of actions -- a plan --
that transforms the world from some initial state to a goal state. In 1992 Kautz
and Selman \cite{Kautz1992} proposed to encode planning as satisfiability. In
their encoding the problem of finding a plan of length $i$ (\textit{i.e.,} the
\emph{makespan}) is translated into a Boolean formula $F_i$ that is satisfiable
if a plan of length $i$ \emph{or less} exists. In later encodings multiple
actions can be executed \emph{in parallel} allowing longer plans to be found by
solving smaller formulas \cite{Rintanen2006, Rintanen2007, Balyo2013}.

Finding the minimal makespan $i$ for which $F_i$ is satisfiable is important for
SAT-based planning in general and the generation of this benchmark set in
particular. The minimal makespan depends on the planning task and the used
encoding. The hardest formulas that a SAT-based planner has to solve are usually
the last unsatisfiable $F_i$ before the next higher makespan $i$ is satisfiable
\cite{Rintanen2006}. Therefore we preferably pick the last unsatisfiable
makespan for each planning task to generate unsatisfiable instances. For
planning tasks where this makespan cannot be determined with available
computational resources, we use a \emph{sequential} encoding, where the minimal
makespan equals the length of the shortest valid plan. Together with known
bounds\footnote{Bounds on plan length are available for some planning tasks from
  the \emph{optimal track} that have \emph{unit cost} actions.} on the optimal
plan length we can generate SAT formulas with predetermined satisfiability for
hard planning problems.

@@

% - 86 out of 100 biggest (number of clauses) are planning
% - 109 / 200 in the planning track are not solved by any solver
% - 186 / 600 in total
% - all 200 above 90\% binary clauses and half have above 99%
The instances of the Planning track are big compared to the main track
instances. Using the number of clauses as a metric; out of the $100$ biggest
instances of both tracks
% of the main-track
$86$ belong to the Planning track. The large size can mainly be attributed to
binary clauses. On average, more than $98$\% of the clauses are binary for
planning instances. The average for the Main track instances is below $60$\%.

@@

@@

A greedy set cover of the solved instances by solvers of the Main track is presented in Table~\ref{tab:gscMain}. 
In the first iteration, the solver which solved the highest number of instances is selected; in our case it was \solver{Kissat-sat} with 264 instances as we know already from Table~\ref{tab:results}.
Given those 264 instances as already covered, \solver{CaDiCaL-alluip} was the best in further contributing to the set by additional $22$ instances in iteration $2$.
We can see that the further iterations tend to add very little,
with the final four iterations adding one instance each. Note that each solver
that managed to solve an instance uniquely (i.e.~being the only solver to do so)
is showing up in the table. This is, for example the case of the last listed  
\solver{exp-V-LGB-MLD-CBT-DL}.
Highlighting unique solutions is a property of the greedy set cover metric
which we believe is congruent with documenting contributions to the state of the art.

@@

\subsection{Similarity of Solvers}
To investigate how similar the solvers of the Main track are to each other, we
define a similarity metric based on the measured runtimes. For that we first
remove $84$ benchmarks that have not been solved by any solver. For the rest of
the benchmarks we use the PAR-2 score,
% based on the \num{5000}-second timeout
i.\,e., for each solver we set a score of \num{10000} for unsolved instances.
This leaves us with $316$ floats
$S_{1},\dots,S_{n}$ for each solver $S$. The \emph{similarity} of two solvers $A$
and $B$, normalized to the interval $[0,1]$ is defined
as:
% by Equation~\ref{eq-similarity}.

% \begin{equation}
% \label{eq-similarity}
\[
	similarity(A, B) =\quad 1 - \frac{\sum{|A_{i} - B_{i}|}}{316 \cdot \num{10000}}
\]
% \end{equation}
We calculated the \emph{similarity} between the $30$ solvers which achieved the
best PAR-2 score in the main track. The results are displayed in
Figure~\ref{fig-similarity-main} as a \emph{heat map}. Additionally the result
of hierarchically clustering the solvers based on their similarity is
illustrated as a \emph{dendrogram}. The height at which two solvers or clusters
are joined reflects how similar they are. For example enabling the \emph{trail
  saving} in \solver{CaDiCaL-alluip} has no impact on the runtime, resulting in
a similarity above \num{0.999}. Thus the two solvers are joined low in the
dendrogram.
A similar visualization was used to visualize the results of the SAT Race \cite{Balyo:2015:SATRace}.

One big cluster we can identify are the \solver{Maple}-decedents to the right,
all of which are modifications of the winners of the SAT Competition 2018 and
the SAT Race 2019.
With the exception of \solver{mergesat} and
maybe \solver{DurianSat} we observe a high similarity within the cluster.
% * 0.9645726159423101, 'exp_V_MLD_CBT_DL', 'Scavel' second highest similarity of all (not 2 configurations of the same solver)
% ** both based on MapleLCMDistChronoBT-dl-v2.2
% ** suggesting that the changes they made are either very similar or both do not have a high impact on the performance of the base solver in this test set.

% The two configurations of the \solver{Relaxed} by \emph{Zhang} and \emph{Cai}
use the same code base as a lot of the solvers in the Maple-Cluster but their
overall performance is better and closer to the CMS-Cluster.

The three CMS versions differ in their implementation of stochastic local search
(SLS) but overall have a similar performance.

The last cluster to the left of the heat map is comprised of the solvers
originally written by Biere. The three configurations of \solver{Kissat} form a
sub-cluster, even though the configurations are fairly distinct from each other.
% , specialized for satisfiable and unsatisfiable instances.
% One targets satisfiable instances, while the other is focused on unsatisfiable
% instances and a third alternating between them.

The other sub-cluster consists of modifications to the 2019 version of
\solver{CaDiCaL}. The new version CaDiCaL-2020 has a worse overall performance
than the solvers in this cluster and a comparatively high similarty to the
Maple-Cluster.

\begin{figure}[tp]
    \centering
    \includegraphics[width=\linewidth]{similarity/cross.pdf}
    \caption{Heat-map and dendrogram (top) based on the runtime similarity of the solvers participating in the main track.
      Darker regions mean that the solvers are more similar.
      A more precise relation between color and similarity-value together with a histogram of the values that appear is given at the bottom.
    }
    \label{fig-similarity-main}
\end{figure}

% \subsection{Solver Ranking Robustness}
\subsection{Influence of Benchmark Selection on Solver Ranking}
The main result of the competition is the ranking of the solvers participating
in the main track, according to their average PAR-2 score. To evaluate the impact
of benchmark selection on the solver ranking, we follow the experiments
described in the tool suite
\texttt{benchfeature}\cite{benchfeature}.
% \footnote{\url{http://fmv.jku.at/benchfeature/}} developed
% by Fazekas, Kaufmann and Biere.
% performed the following
% experiment:

We use random sampling to select a subset of benchmarks used in the Main track.
We start with $316$ benchmarks that have been solved by at least one solver in
the Main track and remove a number of benchmarks randomly.

The solvers are assigned a new rank in ascending order of their PAR-2 score on
each random sample. Note that we never encounter a tie. This ranking can be
seen as an estimate of the original ranking.
If even relatively small random samples result in a good estimate, we can draw a
positive conclusion about the robustness of the ranking.
% We draw a positive conclusion about the robustness of the original ranking, if
% even the resulting estimate from relatively small random samples
% are good approximations.
To determine how similar the estimate is to the original ranking, we calculate
the \emph{Spearman's rank correlation coefficient} of the two rankings. For each
possible subset size (1--316) we generate 50 random samples.
% and compute the correlation coefficient.

Spearman's rank correlation coefficient
for two rankings $r_{1}$ and $r_{2}$
is defined by the following equation:
% \begin{equation}
% \label{eq-spearman}
\[
    \rho = 1 - \frac{6\cdot\sum_{s}{(r_{1}(s) - r_{2}(s))^2}}{n(n^2-1)},
\]
% \end{equation}
where $n$ is the number of solvers in the Main track ($48$), and the rankings
$r_{\{1,2\}}(s)$ map a solver $s$ to its rank, \emph{i.e.,} $1$ for the best
performing solver, whereas $48$ is assigned to the solver
with the highest average PAR-2 score.

The coefficient $\rho$ is in the interval $[-1, 1]$, where a rank correlation of
$1$ means that the two rankings are equal and $-1$ means that one ranking is the
reverse of the other. To give a better intuition for $\rho$, we list a few
modifications to the original ranking together with the resulting rank
correlation coefficient: The smallest change we can make is to switch the rank
of two adjacent solvers, resulting in a high rank correlation $\rho = 0.9999$.
Several small changes also result in a high rank correlation; repeating the same
modification as before $n/2$ times to switch every two adjacent solvers in rank
still gives a value of $\rho = 0.9974$. On the other hand, switching the highest
ranked solver (\solver{Kissat-sat}) with the lowest ranked solver, results in a
rank correlation of $\rho = 0.7602$. Moving \solver{Kissat-sat} to the bottom of
the ranking and moving every other solver up one rank gives a higher
$\rho = 0.8776$. Doing the same to all three \solver{Kissat} configurations
(with ranks 1, 2 and 11, now 46, 47 and 48) results in $\rho = 0.6839$.

The mean and standard deviation of the computed correlation coefficients are
depicted in Figure~\ref{fig:sampleAll}.

\begin{figure}[h]
  \label{fig:sampleAll}
  \includegraphics[width=\textwidth]{stability/ALL_random_smpling_correlations.png}
  \caption{Mean and standard deviation of rank correlation under removal of random instances, 50 samples for each size}
\end{figure}

The rank correlation is high even for relatively small samples.
The average rank correlation drops below $0.99$ only after randomly removing at least $95$ benchmarks, which is $30\%$ of the considered benchmark set. Accordingly, removing fewer benchmarks randomly has almost no effect on the ranking of the solvers.
Furthermore, removing fewer than $200$ ($63\%$) benchmarks, results in an average rank correlation above $0.96$

This suggest that the influence of the random selection in Algorithm~\ref{algo:select} on the solver ranking is limited.
% \ref{sec:byob}

The collected data cannot show, whether all of the benchmarks originally
submitted by the solver authors, have a systematic bias. Since each newly
submitted benchmark family originates from a different domain and is often the
result of current research, we can assume that the submitted families together
are representative. Figure~\ref{fig:sampleFamily} shows the rank correlation
coefficient resulting from removing a complete benchmark family instead of a
random subset. As expected removing a non random subset can have a higher impact
on the ranking even if it is small. Removing the 13 benchmarks from the
\emph{hgen}-family results in a (still high) rank correlation of $0.9895$. Additionally, the ranking of the top five solvers stays the same.

\begin{figure}[h]
  \label{fig:sampleFamily}
  \includegraphics[width=\textwidth]{stability/fam_leave_one_out_corr.png}
  \caption{Change of rank correlation under removal of individual benchmark families}
\end{figure}
